---
title: "SCRX Mouse 1 Report"
author: "Julian Rees"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xlsxjars)
library(rJava)
```

## Summary
55 NOD/SCID mice were received from LBNL Animal Facility on Monday, 2/13.  ^225^Ac was received on Wednesday, 2/16.  On Thursday (2/17, **Day 0**) **55 mice** were weighed and injected IV with the preparatory dose of IgG at 10 \mu g / g.  One mouse expired immediately following injection, likely due to an air embolism. 

After sterile filtering, the amount of IgG.MB1 supplied by Stemcentrx was insufficient to inject all mice. The mice receiving the lowest dose of ^225^Ac and ^177^Lu, as well as the mice receiving the chelator and our backup "*sentinel*" mice (groups E, I, and K) were injected with an analogous solution of IgG obtained from Sigma. 

On Friday (2/18, **Day 1**) **50 mice** were injected with their predetermined dose (25 $\times$ ^225^Ac, 20 $\times$ ^177^Lu, 5 $\times$ chelator).  The study group assignments are given below. 

Group | Agent | Dose (nCi / 25 g) | IgG Source
------|---------|------ | --
A | ^225^Ac | 180 | SCRX
B | ^225^Ac | 60  | SCRX
C | ^225^Ac | 20  | SCRX
D | ^225^Ac | 6.7 | SCRX
E | ^225^Ac | 2.2 | Sigma

Group | Agent | Dose (uCi / 25 g) | IgG Source
------|---------|------ | --
F | ^177^Lu | 120 | SCRX
G | ^177^Lu | 40  | SCRX
H | ^177^Lu | 13.3 | SCRX
I | ^177^Lu | 4.4 | Sigma

Group | Agent | Dose (ug / g) | IgG Source
------|---------|------ | --
K | Chelator | ?? | Sigma
None | None |  None  | Sigma




```{r, echo=FALSE, warning=FALSE}

# this is the processing code here!  reprints of any components will come below
#library(drc)
#library(drfit)
library(ggplot2)
library(reshape2)
library(xlsx)
library(plyr)

mice <- read.xlsx("../Data/testMice.xlsx", header = TRUE, sheetIndex = 1)
mice <- mice[,-1]
mice <- mice[, -ncol(mice)]

# remove premature casualties
mice <- mice[-36,]

# calculate the doses that each mouse received based on the dosing sheet
day0dose <- cut(mice$day0weight, breaks = 15.8+1.5*0:7, 
                labels = 0.1+1:7/100)
day1dose <- cut(mice$day1weight, breaks = 15.8+1.5*0:7, 
                labels = 0.1+1:7/100)
mice_dose <- cbind(mice[,1:3], day0dose, day1dose)

# drop day 0 and modify col names for day to be numeric
mice <- mice[,-2]
colnames(mice) <- c('group', '1', '3', '6')

# create factor for dose radionuclide
agent <- cut(as.numeric(mice$group), breaks = c(0,5,9,10,11),
             labels = c('Ac225','Lu177','Chelator','Sentinal'))
mice <- cbind(mice, agent)

# melt the data to plot 
mmice <- melt(mice, id = c("group", "agent"))

fig_weight <- ggplot(mmice, aes(x = variable, y = value, by = group)) + 
  geom_boxplot(aes(fill = group)) + 
  #geom_line(aes(x = as.numeric(as.character(variable)), color = group)) +
  theme_bw() +
  labs(x = "Day", y = "Weight", fill = "Group") +
  scale_x_discrete(limits=1:6) +
  facet_wrap(~agent)
#fig_weight

# replot as a change in weight, take difference and melt
dmice <- mice
for(i in seq(ncol(dmice)-2)){
  dmice[,i+1] = dmice[,i+1] - mice[,2]
}
mdmice <- melt(dmice, id = c("group", "agent"))

fig_weight_diff <- ggplot(mdmice, aes(x = variable, y = value, by = group)) + 
  geom_boxplot(aes(fill = group, outlier.fill = group)) + 
  #geom_line(aes(x = as.numeric(as.character(variable)), color = group, group = group)) +
  theme_bw() +
  labs(x = "Day", y = "Weight Change", fill = "Group") +
  scale_x_discrete(limits=1:6) +
  facet_wrap(~agent)
#fig_weight_diff

# count up the NA values (dead mice)

survivingmice2 <- ddply(mmice, .(group, variable, agent), summarize, 
      alive = is.finite(value))
survivingmice2 <- subset(survivingmice2, alive == TRUE)

fig_alive_count <- ggplot(survivingmice2, aes(x = variable, 
                                             by = group)) + 
  geom_bar(aes(fill = group), position = position_dodge(width = 1.1)) + 
  labs(x = "Day", y = "Surviving Mice", fill = "Group") +
  theme_bw() +
  scale_x_discrete(limits=1:6) +
  facet_wrap(~agent)
#fig_alive_count
```

## Mouse weights by day
The Day 0 weights for the mice were found to be in the range of **18 - 25 g**, with a single outlier at **16 g**.  The change in mouse weight over time is shown below. The statistical parameters for the box plots are:

* Position = Median
* Height = First to third quartile (25th to 75th percentile)
* Whiskers = Largest deviation within 1.5 $\times$ inter-quartile range (IQR)
* Points = Outlyers beyond 1.5 $\times$ IQR

```{r, echo=FALSE, warning=FALSE}
plot(fig_weight)
```

General trends are hard to observe.  Normalizing each mouse's weight to its Day 1 weight provides a better description of the effects of the administered doses.

```{r, echo=FALSE, warning=FALSE}
plot(fig_weight_diff)
```

The number of mice surviving in each group is tracked below. 

```{r, echo=FALSE, warning=FALSE}
plot(fig_alive_count)
```

